---
title: "Seroprevalence Summary Analyses"
author: "Kate Kennedy"
date: "2023-03-22"
output: html_document
---

```{r 1-setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("00_setup.R")

load("data.RData")
```


## II. Summaries of Exposure Risk Scores


"Central tendencies of exposure risk for each participant"
```{r}
#median, interquartile range, mode, mean, standard deviation, standard error
#geometric mean, geometric standard deviation
er_user <- monthly_er %>%
  group_by(user_id) %>%
  dplyr:::summarise(median_er=median(exposure_risk),
                    IQR_er=IQR(exposure_risk),
                    mean_er=mean(exposure_risk),
                    sd_er=sd(exposure_risk),
                    se_er=se(exposure_risk),
                    geomean_er=exp(mean(log(exposure_risk))),
                    geosd_er=exp(sd(log(exposure_risk))))

```

"Central tendencies of exposure risk for each Month"
```{r}
er_month <- monthly_er %>%
  group_by(month) %>%
  dplyr:::summarise(median_er=median(exposure_risk),
                    IQR_er=IQR(exposure_risk),
                    mean_er=mean(exposure_risk),
                      sd_er=sd(exposure_risk),
                    se_er=se(exposure_risk),
                    geomean_er=exp(mean(log(exposure_risk))),
                    geosd_er=exp(sd(log(exposure_risk))))
```

"Central Tendencies of Exposure Risk for each Month by age Group"
```{r}
er_month_age <- monthly_er %>%
  group_by(month, age_group) %>%
  dplyr:::summarise(median_er=median(exposure_risk),
                    mean_er=mean(exposure_risk),
                    sd_er=sd(exposure_risk),
                    se_er=se(exposure_risk),
                    geomean_er=exp(mean(log(exposure_risk))),
                    geosd_er=exp(sd(log(exposure_risk))))
```
"Central Tendencies of Exposure Risk for each Month by Health Group"
```{r}
er_month_health <- monthly_er %>%
  group_by(month, health) %>%
  dplyr:::summarise(median_er=median(exposure_risk),
                    mean_er=mean(exposure_risk),
                    sd_er=sd(exposure_risk),
                    se_er=se(exposure_risk),
                    geomean_er=exp(mean(log(exposure_risk))),
                    geosd_er=exp(sd(log(exposure_risk))))
```

## III. Subsetting data for demographic analysis and visualizations
"Subset variables of interest"
```{r}
demo <- final_baseline %>%
  dplyr::select(user_id, age, sex, health, age_group, survey)


#Merge Exposure and Demographic Data
er_demo <- er_user %>%
  left_join(demo)
```
###Fig1A and B
```{r}
ill_types <- ill_final %>%
  dplyr::select(user_id, age_group, diagnosis, severity, q0040, q0080, q0360, q0370, q0380, q0390)%>%
  tidyr::replace_na(list(q0040 = 0, q0080 =0, q0360 = 0, q0370 = 0, q0380 = 0, q0390 = 0))%>%
  mutate(illness = if_else(is.na(diagnosis), "Other Symptoms", 
                           if_else(diagnosis == "COVID-19", "COVID-19", 
                                   if_else(diagnosis %in% c("ILI", "Influenza", "Pneumonia"), "Other ILI", "Other Symptoms"))))%>%
  mutate(illness = factor(illness, levels = c("COVID-19", "Other ILI", "Other Symptoms")))
```

```{r}
participants <- final_baseline %>% 
  select(user_id, age, age_group)

ill_all <- ill_types %>%
  group_by(user_id, illness)%>%
  count()%>%
  pivot_wider(names_from = illness, values_from = n)%>%
  full_join(participants)%>%
  tidyr::replace_na(list(`Other Symptoms` = 0, `COVID-19` = 0, `Other ILI` = 0))%>%
  mutate(age_group = factor(age_group, levels = c("Younger", "Middle-aged", "Older"), 
                            labels = c("≤39", "40-59", "≥60")))%>%
  ungroup()

ill_all$total <- rowSums(ill_all[,c("Other Symptoms","Other ILI","COVID-19")])

#colnames(ill_all)

ill_all %>%
  group_by(age_group)%>%
  summarise(min = min(total),
            mean = mean(total),
            median = median(total), 
            max = max(total))

ill_all %>%
  summary()
```


```{r}
lm <- lm(total ~ age_group , data = ill_all)
anova(lm)
em <- emmeans(lm, pairwise~age_group)
pairs(em)
```

```{r}
ill_types_count = ill_types %>% 
  group_by(illness) %>% 
  count() %>% 
  ungroup()%>% 
  arrange(desc(illness)) %>%
  mutate(percentage = round(n/sum(n),3)*100,
         lab.pos = cumsum(percentage)-.5*percentage)

fig1a <- ill_types_count %>%
  ggplot(aes(x = 1, y = percentage, fill = illness))+
  geom_bar(stat = "identity", colour = "black")+
  coord_polar("y", start = 200) +
  geom_text(aes(label = paste(percentage,"%", sep = "")), 
            position = position_stack(vjust = 0.5),
            col = "black")+
  theme_void() +
  theme(legend.title = element_blank(),
        legend.position = "none")+
  scale_fill_manual(values = c("pink", "lightyellow", "lightgrey"))+
  xlim(.2,1.5)
fig1a

ggsave(here::here("figures", "fig1a.svg"))
```

```{r}
infer::chisq_test(ill_types, illness ~ age_group)
```


```{r}
ill_types_count_2 = ill_types %>% 
  group_by(illness, age_group) %>% 
  count() %>% 
  ungroup()%>% 
  arrange(desc(illness)) %>%
  group_by(illness)%>%
  mutate(percentage = round(n/sum(n),4)*100,
         lab.pos = cumsum(percentage)-.5*percentage)

fig1b <- ill_types_count_2 %>%
  ggplot(aes(x = 1, y = percentage, fill = age_group))+
  geom_bar(stat = "identity", colour = "black")+
  coord_polar("y", start = 200) +
  geom_text(aes(label = paste(percentage,"%", sep = "")), 
            position = position_stack(vjust = 0.5),
            col = "black")+
  theme_void() +
  theme(legend.position = "none",
        strip.text = element_blank(),
        panel.spacing=unit(-0.5, "cm"))+
  scale_fill_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  xlim(.2,1.5)+
  facet_grid(~illness)+
  guides(fill=guide_legend(title="Age group"))
fig1b

ggsave(here::here("figures", "fig1b.svg"))
```


```{r}
ill_final %>%
  group_by(age_group)%>%
  dplyr::summarise(n = n(),
            min = fivenum(symp_sum)[1],
            Q1 = fivenum(symp_sum)[2],
            median = fivenum(symp_sum)[3],
            Q3 = fivenum(symp_sum)[4],
            max = fivenum(symp_sum)[5])

ill_final %>%
  group_by(age_group)%>%
  dplyr::summarise(n = n(),
            min = fivenum(q0270)[1],
            Q1 = fivenum(q0270)[2],
            median = fivenum(q0270)[3],
            Q3 = fivenum(q0270)[4],
            max = fivenum(q0270)[5])

n_distinct(ill_final$user_id)
```


```{r}
ill_final %>%
  group_by(health)%>%
  dplyr::summarise(n = n(),
            min = fivenum(symp_sum)[1],
            Q1 = fivenum(symp_sum)[2],
            median = fivenum(symp_sum)[3],
            Q3 = fivenum(symp_sum)[4],
            max = fivenum(symp_sum)[5])

ill_final %>%
  group_by(health)%>%
  dplyr::summarise(n = n(),
            min = fivenum(q0270)[1],
            Q1 = fivenum(q0270)[2],
            median = fivenum(q0270)[3],
            Q3 = fivenum(q0270)[4],
            max = fivenum(q0270)[5])

```

###Fig1C
```{r}
####NOTE#####
#need to change significance asterisks in figures to account for the fact that they aren't accounting for the effect of user_id
lm <- lmer(symp_sum ~ age_group + (1|user_id), data = ill_final)
anova(lm)
em <- emmeans(lm, pairwise~age_group)
pairs(em)

lm <- lmer(symp_sum ~ health + (1|user_id), data = ill_final)
anova(lm)

```


```{r}
fig1c <- ill_final %>%
  ggplot()+
  aes(
    x = age_group,
    y = symp_sum
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                corral = "none",
                corral.width = 0.6,
                aes(colour = age_group),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 1,
    aes(fill = age_group),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               #position = position_dodge(width=0.5), 
               geom = "pointrange")+
  labs(y = "Severity Score", x = "")+
  theme(legend.position = "none")+
  scale_colour_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  scale_fill_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  stat_compare_means(comparisons = list(c("≤39", "≥60")), 
  label = "p.signif", method = "t.test", tip.length = 0)+
  guides(colour=guide_legend(title="Age group"))

fig1c
  
ggsave(here::here("figures", "fig1c.svg"))
```

```{r}
fig1d <- ill_final %>%
  ggplot()+
  aes(
    y = symp_sum,
    x = health,
    colour = health
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                corral = "none",
                corral.width = 0.6,
                aes(colour = health),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 0.5,
    aes(fill = health),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = position_dodge(width=0.5), 
               geom = "pointrange", 
               colour = "black")+
  theme(
    legend.position = "none"
  )+
  labs(y = "Severity Score", x = "")+
  scale_colour_manual(values = c("grey70", "grey10"))+
  scale_fill_manual(values = c("grey70", "grey10"))+
  stat_compare_means(comparisons = list(c("Healthy", "Pre-existing Conditions")), 
  label = "p.signif", method = "t.test", tip.length = 0)
fig1d

ggsave(here::here("figures", "fig1d.svg"))
```

###Fig1D

```{r}
lm <- lmer(q0270 ~ age_group + (1|user_id), data = ill_final)
anova(lm)
em <- emmeans(lm, pairwise~age_group)
pairs(em)

lm <- lmer(q0270 ~ health + (1|user_id), data = ill_final)
anova(lm)
```

```{r}
fig1e <- ill_final %>%
  #filter(q0270 > 0)%>%
  ggplot()+
  aes(
    x = age_group,
    y = q0270
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                corral = "wrap",
                #cex = 1,
                corral.width = 0.6,
                aes(colour = age_group),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 1,
    aes(fill = age_group),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               #position = position_dodge(width=0.5), 
               geom = "pointrange")+
  theme(legend.position = "none")+
  labs(y = "Days of Missed Activities", x = "")+
  scale_colour_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  scale_fill_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  stat_compare_means(comparisons = list(c("≤39", "≥60")), 
  label = "p.signif", method = "t.test", tip.length = 0)+
  #scale_y_continuous(trans = pseudo_log_trans())+
  scale_y_sqrt()+
  guides(colour=guide_legend(title="Age group"))
fig1e
  
ggsave(here::here("figures", "fig1e.svg"))
```

```{r}
fig1f <- ill_final %>%
  #filter(q0270 > 0)%>%
  ggplot(aes(
    x = health,
    y = q0270
  ))+
  geom_beeswarm(side = 1L,
                size = 1,
                corral = "wrap",
                corral.width = 0.6,
                aes(colour = health),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 0.5,
    aes(fill = health),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl,
               fun.args = list(mult = 1),
               geom = "pointrange")+
  labs(y = "Days of Missed Activities", x = "")+
  theme(legend.position = "none")+
  scale_colour_manual(values = c("grey70", "grey20"))+
  scale_fill_manual(values = c("grey70", "grey20"))+
  scale_y_sqrt()+
  stat_compare_means(comparisons = list(c("Healthy", "Pre-existing Conditions")), 
  label = "p.signif", method = "t.test", tip.length = 0)
fig1f

ggsave(here::here("figures", "fig1f.svg"))
```

```{r}
fig1ab <- cowplot::plot_grid(fig1a, fig1b, 
                   labels = c("a", "b"),
                   rel_widths = c(1,2),
                   nrow = 1)

fig1cf <- cowplot::plot_grid(fig1c, fig1d, fig1e, fig1f, 
                   labels = c("c", "d", "e", "f"),
                   nrow = 2)

cowplot::plot_grid(fig1ab, fig1cf, nrow = 2, rel_heights = c(1,2))

ggsave(here::here("figures", "fig1.svg"), width = 8, height = 7 , units = "in")
```



```{r}
ill_counts <- ill_final %>%
  count(user_id)%>%
  arrange()
```


## V. Pre/post ER
```{r}
lm <- lmer(exposure_risk ~ pre_post*age_group + (1|user_id), data = pre_post_long)
anova(lm)
em <- emmeans(lm, pairwise~ pre_post | age_group)
pairs(em)

em <- emmeans(lm, pairwise~ age_group | pre_post)
pairs(em)
```

```{r}
lm <- lmer(exposure_risk ~ pre_post*health + (1|user_id), data = pre_post_long)
anova(lm)
em <- emmeans(lm, pairwise~ pre_post | health)
pairs(em)

em <- emmeans(lm, pairwise~ health | pre_post)
pairs(em)

```


```{r}
fig2a <- pre_post_long %>%
  ggplot()+
  aes(
    y = exposure_risk,
    x = pre_post
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                cex=2.5,
                corral = "wrap",
                corral.width = 0.5,
                aes(colour = age_group),
                stroke = 0, 
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 1,
    aes(fill = age_group),
    colour = NA,
    show.legend = FALSE
    )+
  # geom_beeswarm(alpha = 1, 
  #               aes(colour = age_group), 
  #               shape = 1)+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = position_dodge(width=0.5), 
               geom = "pointrange")+
  facet_grid(~age_group, switch = "both")+
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 1)
  )+
  labs(y = "Exposure Risk", x = "")+
  scale_colour_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  scale_fill_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  stat_compare_means(comparisons = list(c("pre", "post")), 
  label = "p.signif", method = "t.test", paired=T, tip.length = 0)
fig2a

#ggsave(here::here("figures", "fig2a.svg"))
```

```{r}
lm <- lm(change_er ~ age_group, data = pre_post_er)
anova(lm)
em <- emmeans(lm, pairwise~ age_group)
pairs(em)

lm <- lm(change_er ~ health, data = pre_post_er)
anova(lm)

```

```{r}
my_comparisons = list(c("Younger", "Middle-aged"), 
                      c("Middle-aged", "Older"),
                      c("Younger", "Older"))

fig2b <- pre_post_er %>%
  ggplot()+
  aes(
    y = abs(change_er),
    x = age_group
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                cex=1.2,
                corral = "wrap",
                corral.width = 0.5,
                aes(colour = age_group),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 1,
    aes(fill = age_group),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = position_dodge(width=0.5), 
               geom = "pointrange")+
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 1)
  )+
  labs(y = "Decrease in Exposure Risk", x = "Age Group")+
  scale_colour_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  scale_fill_manual(values = c("#a2cafa", "#55a0fd", "#067e97"))+
  stat_compare_means(aes(label = after_stat(p.signif)), 
                     method = "t.test", 
                     ref.group = "≤39")
fig2b

#ggsave(here::here("figures", "fig2a_diff.svg"))
```

```{r}
fig2c <- pre_post_long %>%
  ggplot()+
  aes(
    y = exposure_risk,
    x = pre_post
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                cex=2,
                corral = "wrap",
                corral.width = 0.6,
                aes(colour = health),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 0.5,
    aes(fill = health),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = position_dodge(width=0.5), 
               geom = "pointrange")+
  facet_grid(~health, switch = "both")+
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 1)
  )+
  labs(y = "Exposure Risk", x = "")+
  scale_colour_manual(values = c("grey70", "grey20"))+
  scale_fill_manual(values = c("grey70", "grey20"))+
  stat_compare_means(comparisons = list(c("pre", "post")), 
  label = "p.signif", method = "t.test", paired=T, tip.length = 0)
fig2c

#ggsave(here::here("figures", "fig2b.svg"))
```

```{r}
fig2d <- pre_post_er %>%
  ggplot()+
  aes(
    y = abs(change_er),
    x = health
  )+
  geom_beeswarm(side = 1L,
                size = 1,
                cex=1,
                corral = "wrap",
                corral.width = 0.6,
                aes(colour = health),
                stroke = 0,
                alpha = 0.5)+
  gghalves::geom_half_violin(
    side = "l",
    alpha = 0.5,
    aes(fill = health),
    colour = NA,
    show.legend = FALSE
    )+
  stat_summary(fun.data = mean_sdl, 
               fun.args = list(mult = 1),
               position = position_dodge(width=0.5), 
               geom = "pointrange")+
  theme(
    legend.position = "none",
    strip.background = element_blank(),
    panel.border = element_blank(),
    axis.line = element_line(colour = "black", linewidth = 1)
  )+
  scale_colour_manual(values = c("grey70", "grey20"))+
  scale_fill_manual(values = c("grey70", "grey20"))+
  labs(y = "Decrease in Exposure Risk", x = "")+
  stat_compare_means(comparisons = list(c("Healthy", "Pre-existing Conditions")), 
  label = "p.signif", method = "t.test", tip.length = 0)
fig2d

#ggsave(here::here("figures", "fig2b_diff.svg"))
```

```{r}
cowplot::plot_grid(fig2a, fig2b, fig2c, fig2d,
                   labels = c("a", "b", "c", "d"),
                   nrow = 2)

ggsave(here::here("figures", "fig2.svg"), height = 5, width = 7, units = "in")
```

## VI. Data Visualizations with Exposure Risk and Stringency Index

```{r}
"Overlaid Mean SI and Mean ER by Month"

#Extract mean and standard error values for stringency and exposure
er_si_month <- merge(ox_si_month, er_month, by = "month")
er_si_month <- er_si_month %>% dplyr::select(month, mean_si, se_si,
                                             mean_er, se_er)

#Long form table for plotting
#multiply exposure risk to bring on to same scale as stringency
er_si_table <-  melt(setDT(er_si_month), measure = patterns("mean","se"),
     variable.name = 'Measurement', value.name = c('mean', 'se')) %>% 
  mutate(Measurement = ifelse(Measurement==1, 'Stringency Index', 'Exposure Risk')) %>%
  mutate(mean = ifelse(Measurement == 'Exposure Risk', mean*20, mean))%>%
  mutate(se = ifelse(Measurement == 'Exposure Risk', se*20, se))%>%
  mutate(date = as.Date(lubridate::parse_date_time(month, orders = c("m Y"))))

#Specify Factors for Legend
er_si_table$Measurement <- factor(er_si_table$Measurement, 
                                  levels = c("Stringency Index", "Exposure Risk"))
```


```{r}
#Plot with secondary axis for exposure risk

ggplot(er_si_table, aes(date, mean, colour = Measurement)) +
  geom_point(size=2) +
  geom_pointrange(aes(ymin=mean-se, ymax=mean+se)) +
  scale_colour_manual(values = c("#000063","#CE0000")) +
  scale_y_continuous("Stringency Index",
                     breaks = c(60, 70, 80, 90, 100),
                     sec.axis = sec_axis(~ . / 20, name =
                                           "Exposure Risk")) +
  scale_x_date(date_labels="%b", date_breaks="month") +
  facet_grid(~ year(date), scales="free_x", switch="x") +
  theme(strip.placement = "outside",
        strip.background = element_blank(),
        panel.spacing=unit(0.4,"cm"))+
  theme(panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        strip.text = element_text(size = 14),
        panel.border = element_blank(),
        axis.line = element_line(colour = "black", linewidth = 1), 
        legend.text = element_text(size = 14), 
        legend.title = element_text(size = 16),
        legend.position = "bottom",
        legend.key = element_blank(),
        axis.title.x = element_blank(),
        axis.title = element_text(size = 16, face = 2, colour = "black"), 
        axis.text.x = element_text(size = 12, colour = "black"),
        axis.text.y = element_text(size = 14, colour = "black"),
        plot.title = element_text(color="black", size=18, face="bold", hjust = 0.5, vjust = 2)) + 
  labs(colour = "")


ggsave(here::here("figures", "fig2c.svg"), height = 4.5, width= 8, units = "in")
```